### 2017-10-24 docs/manuscript.tex
updated, preparing for proofs

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Orienting the causal relationship between imprecisely measured traits using GWAS summary data},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Orienting the causal relationship between imprecisely measured traits
using GWAS summary data}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{24 October 2017}

\begin{document}
\maketitle

Gibran Hemani*, Kate Tilling and George Davey Smith

MRC Integrative Epidemiology Unit (IEU) at the University of Bristol,
School of Social and Community Medicine, Bristol, UK

* Correspondence to:
\href{mailto:g.hemani@bristol.ac.uk}{\nolinkurl{g.hemani@bristol.ac.uk}}

\subsubsection{Abstract}\label{abstract}

Inference about the causal structure that induces correlations between
two traits can be achieved by combining genetic associations with a
mediation-based approach, as is done in the causal inference test (CIT).
However, we show that measurement error in the phenotypes can lead to
the CIT inferring the wrong causal direction, and that increasing sample
sizes has the adverse effect of increasing confidence in the wrong
answer. This problem is likely to be general to other mediation-based
approaches. Here we introduce an extension to Mendelian randomisation, a
method that uses genetic associations in an instrumentation framework,
that enables inference of the causal direction between traits, with some
advantages. First, it can be performed using only summary level data
from genome-wide association studies; second, it is less susceptible to
bias in the presence of measurement error or unmeasured confounding. We
apply the method to infer the causal direction between DNA methylation
and gene expression levels. Our results demonstrate that, in general,
DNA methylation is more likely to be the causal factor, but this result
is highly susceptible to bias induced by systematic differences in
measurement error between the platforms, and by horizontal pleiotropy.
We emphasise that, where possible, implementing MR and appropriate
sensitivity analyses alongside other approaches such as CIT is important
to triangulate reliable conclusions about causality.

\subsubsection{Author summary}\label{author-summary}

Understanding the causal relationships between pairs of traits is
crucial for unravelling the causes of disease. To this end, results from
genome-wide association studies are valuable because if a trait is known
to be influenced by a genetic variant then this knowledge can be used to
test the trait's causal influences on other traits and diseases. Here we
discuss scenarios where the nature of the genetic association with the
causal trait can lead existing causal inference methods to give the
wrong direction of causality. We introduce a new method that can be
applied to summary level data and is potentially less susceptible to
problems such as measurement error, and apply it to evaluate the causal
relationships between DNA methylation levels and gene expression. While
our results show that DNA methylation is more likely to be the causal
factor, we point out that is it crucial to acknowledge that systematic
differences in measurement error between the platforms could influence
such conclusions.

\subsection{Introduction}\label{introduction}

Observational measures of the human phenome are growing ever more
abundant, but using these data to make causal inference is notoriously
susceptible to many pitfalls, with basic regression-based techniques
unable to distinguish a true causal association from reverse causation
or confounding {[}1--3{]}. In response to this, the use of genetic
associations to instrument traits has emerged as a technique for
improving the reliability of causal inference in observational data, and
with the coincident rise in genome-wide association studies it is now a
prominent tool that is applied in several different guises {[}3--6{]}.
However, shifting from observational associations to instrumentation
does require more (often untestable) assumptions, and potential pitfalls
remain. One that is often neglected is the influence of non-differential
measurement error on the reliability of causal inference.

Measurement error is the difference between the measured value of a
quantity and its true value. This study focuses specifically on
non-differential measurement error where all strata of a measured
variable have the same error rate, which can manifest as changes in
scale or measurement imprecision (noise). Such variability can arise
through a whole plethora of mechanisms, which are often specific to the
study design and difficult to avoid {[}7,8{]}. Array technology is now
commonly used to obtain high throughput phenotyping at low cost, but
comes with the problem of having imperfect resolution, for instance
methylation levels as measured by the Illumina450k chip are prone to
have some amount of noise around the true value due to imperfect
sensitivity {[}9,10{]}. Relatedly, if the measurement of biological
interest is the methylation level in a T cell, then measurement error of
this value can be introduced by using methylation levels from whole
blood samples because the measured value will be an assay of many cell
types {[}11{]}.

Measurement error will of course arise in other types of data too. For
example when measuring BMI one is typically interested in using this as
a proxy for adiposity, but it is clear that the correlation between BMI
and underlying adiposity is not perfect {[}12{]}, leading to the problem
that phenotypes may be imprecisely defined. A similar problem of
biological misspecification is unavoidable in disease diagnosis, and
measuring behaviour such as smoking or diet is notoriously difficult to
do accurately. Measurement error can also be introduced after the data
have been collected, for example the transformation of non-normal data
for the purpose of statistical analysis will lead to a new variable that
will typically incur both changes in scale and imprecision (noise)
compared to the original variable. The sources of measurement error are
not limited to this list {[}8{]}, and its impact has been explored in
the epidemiological literature extensively {[}13,14{]}. Given the
near-ubiquitous presence of measurement error in phenomic data it is
vital to understand its impact on the tools we use for causal inference.

An established study design that can provide information about causality
is randomisation. Given the hypothesis that trait A (henceforth referred
to as the exposure) is causally related to trait B (henceforth referred
to as the outcome), randomisation can be employed to assess the causal
nature of the association by randomly splitting the sample into two
groups, subjecting one group to the exposure and treating the other as a
control. The association between the exposure and the outcome in this
setting provides a robust estimate of the causal relationship. This
provides the theoretical basis behind randomised control trials, but in
practice randomisation is often difficult or impossible to implement in
an experimental context due to cost, scale or inability to manipulate
the exposure. The principle, however, can be employed in extant
observational data through the use of genetic variants associated with
the exposure (instruments), where the inheritance of an allele serves as
a random lifetime allocation of differential exposure levels
{[}15,16{]}. Two statistical approaches to exploiting the properties of
genetic instruments are widely used: mediation-based approaches and
Mendelian randomisation (MR).

Mediation-based approaches employ genetic instruments (typically single
nucleotide polymorphisms, SNPs) to orient the causal direction between
the exposure and the outcome. If a SNP is associated with an exposure,
and the exposure is associated with some outcome, then it logically
follows that in this simple three-variable scenario the estimated direct
influence of the SNP on the outcome will be zero when conditioning on
the exposure. Here, the exposure completely mediates the association
between the SNP and the outcome, providing information about the causal
influence of the exposure on the outcome. This forms the basis of a
number of methods such as genetical genomics {[}17{]}, the
regression-based causal inference test (CIT) {[}4,18{]}, a structural
equation modelling (SEM) implementation in the NEO software {[}5{]}, and
various other methods including Bayesian approaches {[}6{]}. They have
been employed by a number of recent publications that make causal
inferences in large scale `omics datasets {[}6,19--23{]}.

MR can be applied to the same data - phenotypic measures of the exposure
and the outcome variables and a genetic instrument for the exposure -
but the genetic instrument is employed in a subtly different manner.
Here the SNP is used as a surrogate for the exposure. Assuming the SNP
associates with the outcome only through the exposure, the causal effect
of the exposure on the outcome can be estimated by scaling the
association between the SNP and the outcome by the association between
the SNP and the exposure. Though difficult to test empirically, this
assumption can be relaxed in various ways when multiple instruments are
available for a putative exposure {[}24,25{]} and a number of
sensitivity tests are now available to improve reliability {[}26{]}.
Additionally, if valid genetic instruments are known for both traits of
interest then MR can be performed in both directions (bi-directional
MR), testing the influence of one trait on the other and vice versa, to
infer the causal direction between the two phenotypes {[}27,28{]}.

By utilising genetic instruments in different ways, mediation-based
analysis and MR models have properties that confer some advantages and
some disadvantages for reliable causal inference. In the CIT framework
(described fully in the Methods) for example, the test statistic is
different if you test for the exposure causing the outcome or the
outcome causing the exposure, allowing the researcher to infer the
direction of causality between two variables by performing the test in
both directions and choosing the model with the strongest evidence. The
CIT also has the valuable property of being able to distinguish between
several putative causal graphs that link the traits with the SNP (Figure
1). Such is not the case for MR, where in order to infer the direction
of causality between two traits the instrument must have its most
proximal link with the exposure, associating with the outcome only
through the exposure.

Assuming biological knowledge of genetic associations can be problematic
because if there exists a putative association between two variables,
with the SNP being robustly associated with each, it can be difficult to
determine which of the two variables is subject to the primary effect of
the SNP (i.e.~for which of the two variables is the SNP a valid
instrument? Figure 1). By definition, we expect that if the association
is causal then a SNP for the exposure will be associated with the
outcome, such that if the researcher erroneously uses the SNP as an
instrument for the outcome then they are likely to see an apparently
robust causal association of outcome on exposure. Genome-wide
association studies (GWASs) that identify genetic associations for
complex traits are, by design, hypothesis free and agnostic of genomic
function, and it often takes years of follow up studies to understand
the biological nature of a putative GWAS hit {[}29{]}. If multiple
instruments are available for an hypothesised exposure, which is
increasingly typical for complex traits that are analysed in large GWAS
consortia, then techniques can be applied to mitigate these issues
{[}16{]}. But these techniques cannot always be applied in the case of
determining causal directions between 'omic measures where typically
only one cis-acting SNP is known. For example if a DNA methylation probe
is associated with expression of an adjacent gene, then is a cis-acting
SNP an instrument for the DNA methylation level, or the gene expression
level (Figure 1)?

MR has some important advantages over the mediation-based approaches.
First, the mediation-based approaches require that the exposure, outcome
and instrumental variables are all measured in the same data, whereas
recent extensions to MR circumvent this requirement, allowing causal
inference to be drawn when exposure variables and outcome variables are
measured in different samples {[}30{]}. This has the crucial advantage
of improving statistical power by allowing analysis in much larger
sample sizes, and dramatically expands the breadth of possible
phenotypic relationships that can be evaluated {[}26{]}. Second, the
mediation-based approach of adjusting the outcome for the exposure to
nullify the association between the SNP and the outcome is affected by
unmeasured confounding of the exposure and outcome. This is because
adjusting the outcome by the exposure induces a collider effect between
the SNP and outcome {[}31{]}, and in order to fully abrogate this
association one must also adjust for all (hidden or otherwise)
confounders. MR does not suffer from this problem because it does not
test for association through adjustment. Third, when MR assumptions are
satisfied the method is robust to there being measurement error in the
exposure variable {[}32{]}. Indeed instrumental variable (IV) analysis
was in part initially introduced as a correction for measurement error
in the exposure {[}33{]}, whereas it has been noted that both classic
mediation-based analyses {[}13,14,34,35{]} and mediation-based methods
that use instrumental variables {[}36,37{]} are prone to be unreliable
in its presence.

Using theory and simulations we show how non-differential measurement
error in phenotypes can lead to unreliable causal inference in the
mediation-based CIT method. Though we only examine the CIT method in
detail, we believe that attempting to adjust for mediating variables to
make causal inference is susceptible to problems, which can be
generalised to other mediation-based methods. We then present an
extension to MR that allows researchers to ascertain the causal
direction of an association even when the biology of the instruments are
not fully understood, and also a metric to evaluate the sensitivity of
the result of this extension to measurement error. Finally, to
demonstrate the potential impact of measurement error we apply this
method to infer the direction of causation between DNA methylation
levels and gene expression levels. Our analyses highlight that because
these different causal inference techniques have varying strengths and
weaknesses, triangulation of evidence from as many sources as possible
should be practiced in causal inference {[}38{]}.

\subsection{Model}\label{model}

We model a system whereby some exposure \(x\) has a causal influence
\(\beta_x\) on an outcome \(y\) such that

\[
y = \alpha_x + \beta_x x + \epsilon_x
\]

In addition, the exposure is influenced by a SNP \(g\) with an effect of
\(\beta_g\) such that

\[
x = \alpha_g + \beta_g g + \epsilon_g
\]

The \(\alpha_*\) terms represent intercepts, and henceforth can be
ignored. The \(\epsilon_*\) terms denote random error, assumed
independently and normally distributed with mean zero. Mediation-based
analyses that test whether \(x\) causally relates to \(y\) rely on
evaluating whether the influence of \(g\) on \(y\) can be accounted for
by conditioning on \(x\), such that

\[
cov(g, y - \hat{y}) = 0
\]

where \(\hat{y} = \hat{\beta}_x x\) and assuming no intercept
\(y - \hat{y} = \epsilon_x\). MR analysis estimates the causal influence
of \(x\) on \(y\) by using the instrument as a proxy for \(x\), such
that

\[
\begin{aligned}
\hat{x} & = \hat{\beta}_g g \\
y       & = \beta_{MR}\hat{x} + \epsilon_{MR}
\end{aligned}
\]

where \(\beta_{MR} \neq 0\) denotes the existence of causality, and
\(\beta_{MR}\) is an estimate of the causal effect.

Measurement error of an exposure can be modeled as a transformation of
the true value (\(x\)) that leads to the observed value, \(x_o = f(x)\).
For example, following Pierce and VanderWeele {[}32{]} we can define

\[
f(x) = \alpha_{mx} + \beta_{mx} x + \epsilon_{mx}
\]

where \(\alpha_{mx}\) and \(\beta_{mx}\) influence the error in the
measurement of \(x\) by altering its scale, and \(\epsilon_{mx}\)
represents the imprecision (or noise) in the measurement of \(x\).
Measurement imprecision can represent imprecise measurement due to
limits on sensitivity of measuring equipment, or arise because of
phenotypes being imprecisely defined. The same model of measurement
error can be applied to the outcome variable \(y\).

In this study we assume there is no measurement error in the SNP. Common
genetic variants are typically less susceptible to measurement error due
to strict quality control procedures prior to genome wide association
studies. Any non-differential measurement error that might be present
(either because the SNP is poorly typed or because the SNP is not in
complete linkage disequilibrium with the causal variant) will reduce
power in MR but will not incur bias {[}3,13,32{]}. We also assume that
measurement error in the exposure and the outcome are uncorrelated.

\subsection{Methods}\label{methods}

\subsubsection{CIT test}\label{cit-test}

First we describe how the CIT method {[}4{]} is implemented in the R
package \emph{R/cit} {[}18{]}. The methodology of the CIT is as follows.
Assume an exposure \(x\) is instrumented by a SNP \(g\), and the
exposure \(x\) causes an outcome \(y\), as described above. The
following tests are then performed:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \(H_0: cov(g, y) = 0; H_1: cov(g, y) \neq 0\); \emph{the SNP
  associates with the outcome}
\item
  \(H_0: cov(g, x|y) = 0; H_1: cov(g, x|y) \neq 0\); \emph{the SNP
  associates with the exposure conditional on the outcome}
\item
  \(H_0: cov(x, y|g) = 0; H_1: cov(x, y|g) \neq 0\); \emph{the exposure
  associates with the outcome conditional on the SNP}
\item
  \(H_0: cov(g, y|x) \neq 0; H_1: cov(g, y|x) = 0\); \emph{the SNP is
  independent of the outcome conditional on the exposure}
\end{enumerate}

The term in the 4th test can be rewritten as a
\(cov(g, y|x) = cov(g, y - \hat{y})\) where
\(y - \hat{y} = y - \hat{\alpha}_g + \hat{\beta}_g x\) is the residual
of \(y\) after adjusting for \(x\), and \(x\) is assumed to mediate the
association between the SNP and the outcome. The condition in the 4th
test is formulated as an equivalence testing problem that is estimated
using simulations, comparing the estimate from the data against
empirically obtained estimates for simulated variables where the
independence model is true (full details are given in {[}4{]}). We note
here that this approach is liable to fail, even when there is a true
causal relationship, when confounders of the exposure and outcome are
present, as these will induce collider bias.

If all four tests reject the null hypothesis then it is inferred that
\(x\) causes \(y\). The CIT measures the strength of causality by
generating an omnibus p-value, \(p_{CIT}\), which is simply the largest
(least extreme) p-value of the four tests, the intuition being that
causal inference is only as strong as the weakest link in the chain of
tests.

Now we describe how we used the CIT method in our simulations. The
\emph{cit.cp} function was used to obtain an omnibus p-value. To infer
the direction of causality using the CIT method, an omnibus p-value
generated by CIT for each of two tests - \(p_{CIT, x \rightarrow y}\),
was estimated for the direction of \(x\) causing \(y\) (Model 1), and
for the direction of \(y\) causing \(x\), \(p_{CIT, y \rightarrow x}\)
(Model 2). The results from each of these methods can then be used in
combination to infer the existance and direction of causality. For some
significance threshold \(\alpha\) there are four possible outcomes from
these two tests, and their interpretations are as follows:

\begin{itemize}
\tightlist
\item
  If \(p_{CIT, x \rightarrow y} < \alpha\) and
  \(p_{CIT, y \rightarrow x} > \alpha\) then model 1 is accepted
\item
  If \(p_{CIT, x \rightarrow y} > \alpha\) and
  \(p_{CIT, y \rightarrow x} < \alpha\) then model 2 is accepted
\item
  If \(p_{CIT, x \rightarrow y} > \alpha\) and
  \(p_{CIT, y \rightarrow x} > \alpha\) then no evidence for a causal
  relationship
\item
  If \(p_{CIT, x \rightarrow y} < \alpha\) and
  \(p_{CIT, y \rightarrow x} < \alpha\) then there is potentially
  confounding (S1 fig) and no call is made.
\end{itemize}

For the purposes of compiling simulation results we use an arbitrary
\(\alpha = 0.05\) value, though we stress that for real analyses it is
not good practice to rely on p-values for making causal inference, nor
is it reliable to depend on arbitrary significance thresholds {[}39{]}.

\subsubsection{MR causal test}\label{mr-causal-test}

Two stage least squares (2SLS) is a commonly used technique for
performing MR when the exposure, outcome and instrument data are all
available in the same sample. A p-value for this test, \(p_{MR}\), was
obtained using the \(systemfit\) function in the R package
\(R/systemfit\) {[}40{]}. Note that the value of \(p_{MR}\) is identical
when using the same genetic variant to instrument the influence of the
exposure \(x\) on the outcome \(y\), or erroneously, instrumenting the
outcome \(y\) on the exposure \(x\).

The method that we will now describe is designed to distinguish between
two models, \(x \rightarrow y\) or \(y \rightarrow x\). Unlike the CIT
framework, this approach cannot infer if the true model is
\(x \leftarrow g \rightarrow y\). We also assume all genetic effects are
additive.

To infer the direction of causality it is desirable to know which of the
variables, \(x\) or \(y\), is being directly influenced by the
instrument \(g\). This can be achieved by assessing which of the two
variables has the biggest absolute correlation with \(g\) (S2 Text),
formalised by testing for a difference in the correlations \(\rho_{gx}\)
and \(\rho_{gy}\) using Steiger's Z-test for correlated correlations
within a population {[}41{]}. It is calculated as

\[
Z = (Z_{gx} - Z_{gy}) \frac{\sqrt{N-3}}{\sqrt{2(1-\rho_{xy})h}}
\]

where Fisher's z-transformation is used to obtain
\(Z_{g*} = \frac{1}{2} \ln \left ( \frac{1+\rho_{g*}}{1-\rho_{g*}} \right )\),

\[
h = \frac{1 - (frm^2)} {1 - rm^2}
\]

where

\[
f = \frac{1 - \rho_{xy}}{2(1 - rm^2)}
\]

and

\[
rm^2 = \frac{1}{2}(\rho_{gx}^2 + \rho_{gy}^2).
\]

The \(Z\) value is interpreted such that

\[
Z \left\{
\begin{array}{ll}
> 0, & x \to y\\
< 0, & y \to x\\
= 0, & x \perp\!\!\!\perp y 
\end{array} \right.
\]

and a p-value, \(p_{Steiger}\) is generated from the \(Z\) value to
indicate the probability of obtaining a difference between correlations
\(\rho_{gx}\) and \(\rho_{gy}\) at least as large as the one observed,
under the null hypothesis that both correlations are identical.

The existence of causality and its direction is inferred based on
combining information from the MR analysis and the Steiger test. The MR
analysis indicates whether there is a potential causal relationship
(\(p_{MR}\)), and the Steiger test indicates the direction (\(sign(Z)\))
of the causal relationship and the confidence of the direction
(\(p_{Steiger}\)). For the purposes of compiling simulation results,
these can be combined using an arbitrary \(\alpha = 0.05\) value:

\begin{itemize}
\tightlist
\item
  If \(p_{Steiger} < \alpha\) and \(p_{MR} < \alpha\) and \(Z > 0\) then
  a causal association for the correct model is accepted,
  \(x \rightarrow y\)
\item
  If \(p_{Steiger} < \alpha\) and \(p_{MR} < \alpha\) and \(Z < 0\) then
  a causal association for the incorrect model is accepted,
  \(y \rightarrow x\)
\item
  Otherwise if \(p_{Steiger} > \alpha\) or \(p_{MR} > \alpha\), neither
  model is accepted
\end{itemize}

Note that the same correlation test approach can be applied to a
two-sample MR setting. Two-sample MR refers to the case where the
SNP-exposure association and SNP-outcome association are calculated in
different samples (e.g.~from publicly available summary statistics
{[}26,30{]}). Here the Steiger test of two independent correlations can
be applied where.

\[
Z = \frac{Z_{gx} - Z_{gy}} { \sqrt{ 1 / (N_{1} - 3) + 1 / (N_{2} - 3) } }
\]

An advantage of using the Steiger test in the two sample context is that
it can compare correlations in independent samples where sample sizes
are different. Steiger test statistics were calculated using the
\emph{r.test} function in the R package \emph{R/psych} {[}42{]}.

The Steiger test assumes that there is a causal relationship between the
two variables, and that the SNP is a valid instrument for one of them.
However it is liable to give incorrect causal directions under some
other circumstances. First, some levels of horizontal pleiotropy, where
the SNP influences the outcome through some pathway other than the
exposure, could induce problems because this is a means by which the
instrument is invalid. Second, some differential values of measurement
error between the exposure and the outcome could lead to incorrect
inference of the causal direction (S2 Text). Third, some levels of
unmeasured confounding between the exposure and the outcome could lead
to inference of the wrong causal direction (S3 Text).

\subsubsection{Causal direction sensitivity analysis for measurement
error}\label{causal-direction-sensitivity-analysis-for-measurement-error}

The Steiger test for inferring if \(x \rightarrow y\) is based on
evaluating \(\rho_{gx} > \rho_{gy}\). However, \(\rho_{gx}\) (or
\(\rho_{gy}\)) are underestimated if \(x\) (or \(y\)) are measured
imprecisely. If, for example, \(x\) has lower measurement precision than
\(y\) then we might empirically obtain \(\rho_{g,x_o} < \rho_{g,y_o}\)
because \(\rho_{g,x_o}\) could be underestimated more than
\(\rho_{g,y_o}\).

As we show in S2 Text it is possible to infer the bounds of measurement
error on \(x_o\) or \(y_o\) given known genetic associations. The
maximum measurement imprecision of \(x_o\) is \(\rho_{g,x_o}\), because
it is known that at least that much of the variance has been explained
in \(x_o\) by \(g\). The minimum is 0, denoting perfectly measured trait
values (the same logic applies to \(y_o\)). It is possible to simulate
what the inferred causal direction would be for all values within these
bounds.

To evaluate how reliable, \(R\), the inference of the causal direction
is to potential measurement error in \(x\) and \(y\) we need to predict
the values of \(\rho_{gy} - \rho_{gx}\) for those values of measurement
error. We offer two tools in which to do this. First, the user can
provide values of measurement error for \(x\) and \(y\) and obtain a
revised inference of the causal direction. Second, we integrate over the
entire range of \(\rho_{gy} - \rho_{gx}\) values for possible
measurement error values, assuming that any measurement error value is
equally likely. Across all possible values of measurement error in \(x\)
and \(y\) we find the volume that agrees with the inferred direction of
causality and the volume that disagrees with the inferred direction of
causality, and take the ratio of these two values. A ratio \(R=1\)
indicates that the inferred causal direction is highly sensitive to
measurement error, because equal weight of the measurement error
parameter space supports each direction of causality. In general, the
\(R\) value denotes that the inferred direction of causality is \(R\)
times more likely to be the empirical result than the opposite direction
(S2 Text).

\subsubsection{Simulations}\label{simulations}

Simulations were conducted by creating variables of sample size \(n\)
for the exposure \(x\), the measured values of the exposure \(x_o\), the
outcome \(y\), the measured values of the outcome \(y_o\) and the
instrument \(g\). One of two models are simulated, the ``causal model''
where \(x\) causes \(y\) and \(g\) is an instrument for \(x\); or the
``non-causal model'' where \(g\) influences a confounder \(u\) which in
turn causes both \(x\) and \(y\). Here \(x\) and \(y\) are correlated
but not causally related. Each variable in the causal model was
simulated such that:

\[
\begin{aligned}
g & \sim Binom(2, 0.5) \\
x & = \alpha_g + \beta_g g + \epsilon_g \\
x_o & = \alpha_{mx} + \beta_{mx} x + \epsilon_{mx} \\
y & = \alpha_x + \beta_x x + \epsilon_x \\
y_o & = \alpha_{my} + \beta_{my} y + \epsilon_{my} \\
\end{aligned}
\]

where non-differential measurement error is represented by a noise
(measurement imprecision) term
\(\epsilon_{m*} \sim N(0, \sigma^2_{m*})\), and measurement bias terms
\(\alpha_{m*}\) and \(\beta_{m*}\) for the exposure variable \(x\) and
the outcome variable \(y\). Note that following the first section of the
Results we no longer include the bias terms for simplicity. We have
formulated the non-causal model as:

\[
\begin{aligned}
y & = \alpha_{gy} + \beta_{gy} g + \epsilon_{gy} \\
\end{aligned}
\]

All \(\alpha\) values were set to 0, and \(\beta\) values set to 1.
Normally distributed values of \(\epsilon_*\) were generated such that

\[
\begin{aligned}
cor(g, x)^2 & = 0.1 \\
cor(x, y)^2 & = \{0.2, 0.4, 0.6, 0.8\} \\
\sigma^2_{mx} & = \{0, 0.2, 0.4, 0.6, 0.8, 1\} \\
\sigma^2_{my} & = \{0, 0.2, 0.4, 0.6, 0.8, 1\} \\
n & = \{100, 1000, 10000\}
\end{aligned}
\]

giving a total of 432 combinations of parameters. Simulations using each
of these sets of variables were performed 100 times, and the CIT and MR
methods were applied to each in order to evaluate the causal association
of the simulated variables. Similar patterns of results were obtained
for different values of \(cor(g, x)\).

\subsection{Applied example using two sample
MR}\label{applied-example-using-two-sample-mr}

Two sample MR {[}30{]} was performed using summary statistics for
genetic influences on gene expression and DNA methylation. To do this we
obtained a list of 458 gene expression - DNA methylation associations as
reported in Shakhbazov et al {[}43{]}. These were filtered to be located
on the same chromosome, have robust correlations after correcting for
multiple testing, and to share a SNP that had a robust cis-acting effect
on both the DNA methylation probe and the gene expression probe. Because
only summary statistics were available (effect, standard error, effect
allele, sample size, p-values) for the instrumental SNP on the
methylation and gene expression levels, the Steiger test of two
independent correlations was used to infer the direction of causality
for each of the associations. The Wald ratio test was then used to
estimate the causal effect size for the estimated direction for each
association.

All analysis was performed using the R programming language {[}44{]} and
code is made available at
\url{https://github.com/explodecomputer/causal-directions} and
implemented in the MR-Base (\url{http://wwww.mrbase.org}) platform
{[}26{]}.

\subsection{Results}\label{results}

\subsubsection{Mediation-based causal inference under measurement
error}\label{mediation-based-causal-inference-under-measurement-error}

In the causal inference test (CIT), the 4th condition (see Methods)
employs mediation for causal inference, and can be expressed as
\(cov(g, y - \hat{y}) = 0\), where
\(\hat{y} = \hat{\alpha}_x + \hat{\beta}_x x_o\). When measurement error
in scale and imprecision is introduced, such that \(y_o\) is the
measured value of \(y\), it can be shown using basic covariance
properties (S1 Text) that

\[
\begin{aligned}
cov(g, y - \hat{y}) & = cov(g, y_o) - cov(g, \hat{y}_o)  \\
                    & = \beta_{my} \beta_g \beta_x var(g) - D \beta_{my} \beta_g \beta_x var(g)
\end{aligned}
\]

where

\[
D = \frac{\beta^2_{mx} var(x)} {\beta^2_{mx} var(x) + var(\epsilon_{mx})}
\]

Thus an observational study will find \(cov(g, y_o - \hat{y_o}) = 0\)
when the true model is causal only when \(D = 1\). Therefore, if there
is any measurement error that incurs imprecision in \(x\) (i.e.
\(var(\epsilon_{mx}) \neq 0\)) then there will remain an association
between \(g\) and \(y_o | x_o\), which is in violation of the the 4th
condition of the CIT. Note that scale transformation of \(x\) or \(y\)
without any incurred imprecision is insufficient to lead to a violation
of the test statistic assumptions, and henceforth mention of measurement
error will relate to imprecision unless otherwise stated.

We performed simulations to verify that this problem does arise using
the CIT method. Figure 2 shows that when there is no measurement error
in the exposure or outcome variables (\(\rho_{x, x_o}=\rho_{y, y_o}=1\))
the CIT is reliable in identifying the correct causal direction.
However, as measurement error increases in the exposure variable,
eventually the CIT is more likely to infer a robust causal association
in the wrong direction. Also of concern here is that increasing sample
size does not solve the issue, indeed it only strengthens the apparent
evidence for the incorrect inference.

\subsubsection{Using MR Steiger to infer the direction of
causality}\label{using-mr-steiger-to-infer-the-direction-of-causality}

If we do not know whether the SNP \(g\) has a primary influence on \(x\)
or \(y\) then CIT can attempt to infer the causal direction. Though
bi-directional MR can be used to orient causal directions {[}27{]}, this
requires knowledge of a valid instrument for each trait, and we were
motivated to develop the MR Steiger method that could operate on summary
data to orient the direction of causality using the same conditions as
the CIT, where the underlying biology of a single SNP is not fully
understood. We go on to explore the scenarios in which the method is
likely to return the correct or incorrect causal directions.

We performed simulations to compare the power and type 1 error rates of
MR and CIT in detecting a causal association between simulated variables
under different levels of imprecision simulated in the exposure.
Comparing the performance of methods with different sets of assumptions
can be difficult, but a basic comparison is shown in Figure 3. We
observe that the CIT is more conservative under the null model of no
association owing to the omnibus test statistic comprising several
statistical tests. The FDR using a p-value threshold of 0.05 appears to
be close to zero, whereas for the MR Steiger method the FDR is around
0.05. Using the same p-value thresholds to declare significance in the
non-null simulations, the general trend appears to be that the CIT power
reduces as measurement error in the exposure increases more steeply than
that of the MR Steiger method.

For a particular association, it is of interest to identify the range of
possible measurement error values for which the method will give results
that agree or disagree with the empirically inferred causal direction
(Figure 4a, S2 Text). This metric can be used to evaluate the
reliability of MR Steiger test.

We show that in the presence of measurement imprecision,
\(d = \rho_{x, x_o} - \rho_{x,y}\rho_{y,y_o}\) (S2 Text) determines the
range of parameters around which the MR Steiger test is liable to
provide the wrong direction of causality (\emph{i.e.} if \(d>0\) then
the MR Steiger test is likely to be correct about the causal direction).
Figure 4b shows that when there is no measurement error in \(x\), the MR
Steiger test is unlikely to infer the wrong direction of causality even
if there is measurement error in \(y\). It also shows that in most cases
where \(x\) is measured with error, especially when the causal effect
between \(x\) and \(y\) is not very large, the sensitivity of the MR
Steiger test to measurement error is relatively low.

Unmeasured confounding between the exposure and outcome can also give
rise to problems with the MR Steiger approach (S3 Text). The
relationship between unmeasured confounding and causal orientation is
complex across the parameter space of possible confounding values (S2
Fig). Based on the range of parameter values that we explored, when the
magnitude of the observational variance explained between the exposure
and the outcome is below 0.2 the MR Steiger method is unlikely to return
the incorrect causal direction due to unmeasured confounding.

\subsubsection{Comparison of CIT and MR Steiger for obtaining the
correct direction of
causality}\label{comparison-of-cit-and-mr-steiger-for-obtaining-the-correct-direction-of-causality}

We used simulations to explore the performance of the MR Steiger
approach in comparison to CIT for different levels of measurement error.
The performance was compared in terms of the rate at which evidence of a
causal relationship is obtained for the correct direction of causality,
and the rate at which evidence of a causal relationship is obtained
where the reported direction of causality is incorrect. Simulations were
performed for two models, one for a ``causal model'' where there was a
causal relationship between \(x\) and \(y\); and one for a ``non-causal
model'' where \(x\) and \(y\) were not causally related, but had a
confounded association induced by the SNP \(g\) influencing \(x\) and
\(y\) independently.

Figure 5a shows that, for the ``causal model'', the MR analysis is
indeed liable to infer the wrong direction of causality when \(d < 0\),
and that this erroneous result is more likely to occur with increasing
sample size. However, the CIT is in general more fallible to reporting a
robust causal association for the wrong direction of causality. When
\(d > 0\) we find that in most cases the MR Steiger method has greater
power to obtain evidence for causality than CIT, and always obtains the
correct direction of causality. The CIT, unlike the MR Steiger test, is
able to distinguish the ``non-causal model'' from the ``causal model''
(Methods, Figure 5b), but it is evident that measurement error will
often lead the CIT to identify the causal model as true, when in fact
the underlying model is this non-causal model.

\subsubsection{The causal relationship between gene expression and DNA
methylation
levels}\label{the-causal-relationship-between-gene-expression-and-dna-methylation-levels}

We used the MR Steiger test to infer the direction of causality between
DNA methylation and gene expression levels between 458 putative
associations. We found that the causal direction commonly goes in both
directions (Figure 6a), but assuming no or equal measurement error, DNA
methylation levels were the predominant causal factor
(\(p = 1.3\times 10^{-5}\)). The median reliability (R) of the 458 tests
was 3.92 (5\%-95\% quantiles 1.08 - 37.11). We then went on to predict
the causal directions of the associations for varying levels of
systematic measurement error for the different platforms. Figure 6a
shows that the conclusions about the direction of causality between DNA
methylation and gene expression are very sensitive to measurement error.
We made a strong assumption that either methylation influenced gene
expression or vice versa, but it is certainly possible that the SNP is
solely or additionally influencing some other trait that confounds the
association between gene expression and DNA methylation.

We performed two sample MR {[}30{]} for each association in the
direction of causality inferred by the Stieger test. We observed that
the sign of the MR estimate was generally in the same direction as the
Pearson correlation coefficient reported by Shakhbazov et al {[}43{]}
(Figure 6b). There was a moderate correlation between the absolute
magnitudes of the causal correlation and the observational Pearson
correlation (r = 0.45). Together these inferences suggest that even in
estimating associations between `omic' variables, which are considered
to be low level phenotypes, it is important to use causal inference
methods over observational associations to infer causal effect sizes.

We also observed that for associations where methylation caused gene
expression the causal effect was more likely to be negative than for the
associations where gene expression caused methylation (OR = 0.61 (95\%
CI 0.36 - 1.03), Figure 6c), suggesting that reducing methylation levels
at a controlling CpG typically leads to increased gene expression
levels, consistent with expectation {[}45{]}.

\subsection{Discussion}\label{discussion}

Researchers are often confronted with the problem of making causal
inferences using a statistical framework on observational data. In the
epidemiological literature issues of measurement error in mediation
analysis are relatively well explored {[}46{]}. Our analysis extends
this to related methods such as CIT that are used in predominantly 'omic
data. These methods are indeed susceptible to the same problem as
standard mediation based analysis, and specifically we show that as
measurement error in the (true) exposure variable increases, CIT is
likely to have reduced statistical power, and liable to infer the wrong
direction of causality. We also demonstrate that, though unintuitive,
increasing sample size does not resolve the issue, rather it leads to
more extreme p-values for the model that predicts the wrong direction of
causality.

Under many circumstances a practical solution to this problem is to use
Mendelian randomisation instead of methods such as the CIT or similar
that are based on mediation. Inferring the existence of causality using
Mendelian randomisation is robust in the face of measurement error and,
if the researcher has knowledge about the biology of the instrument
being used in the analysis, can offer a direct solution to the issues
that the CIT faces. This assumption is often reasonable, for example
SNPs are commonly used as instruments when they are found in genes with
known biological relevance for the trait of interest. But on many
occasions, especially in the realm of 'omic data, this is not the case,
and methods based on mediation have been valuable in order to be able to
both ascertain if there is a causal association and to infer the
direction of causality. Here we have described a simple extension to MR
which can be used as an alternative to or in conjunction with mediation
based methods. We show that this method is still liable to measurement
error, but because it has different properties to the CIT it offers
several main advantages. First, it uses a formal statistical framework
to test for the reliability of the assumed direction of causality.
Second, after testing in a comprehensive range of scenarios the MR based
approach is less likely to infer the wrong direction of causality
compared to CIT, while substantially improving power over CIT in the
cases where \(d > 0\).

We demonstrate this new method by evaluating the causal relationships of
458 known associations between DNA methylation and gene expression
levels using summary level data. The inferred causal direction is
heavily influenced by how much measurement error is present in the
different assaying platforms. For example, if DNA methylation measures
typically have higher measuremet error than gene expression measures
then our analysis suggests that DNA methylation levels would be more
often the causal factor in the association. Indeed, previous studies
which have evaluated measurement error in these platforms do support
this position {[}47,48{]}, though making strong conclusions for this
analysis is difficult because measurement error is likely to be study
specific. We also haven't accounted for the influence of winner's curse,
which can inflate estimates of the variance explained by SNPs, with
higher inflation expected amongst lower powered studies. Using p-values
for genetic associations from replication studies will mitigate this
problem.

In our simulations we focused on the simple case of a single instrument
in a single sample setting with a view to making a fair comparison
between MR and the various mediation-based methods available. However,
if there is only a single instrument it is difficult to separate between
the two competing models of \(g\) instrumenting a trait which causes
another trait, and \(g\) having pleiotropic effects on both traits
independently {[}49{]}. Under certain conditions of measurement error
the CIT test can distinguish these models. We also note that it is
straightforward to extend the MR Steiger approach to multiple
instruments, requiring only that the total variance explained by all
instruments be calculated under the assumption that they are
independent. Multiple instruments can indeed help to distinguish between
the causal and pleiotropic models, for example by evaluating the
proportionality of the SNP-exposure and SNP-outcome effects {[}16{]}.
Additionally, if there is at least one instrument for each trait then
bi-directional MR can offer solutions to inferring the causal direction
{[}16,28,50{]}. We restricted the simulations to evaluating the causal
inference between quantitative traits, but it is possible that the
analysis could be extended to binary traits by using the genetic
variance explained on the liability scale, taking into account the
population prevalence {[}51{]}. However, our analysis goes beyond many
previous explorations of measurement error by assessing the impacts of
both imprecision (noise) and linear transformations of the true variable
on causal inference.

Our new method attempts to infer causal directions under the assumption
that horizontal pleiotropy (the influence of the instrument on the
outcome through a mechanism other than the exposure) is not present.
Recent method developments in MR {[}24,25{]} have focused on accounting
for the issues that horizontal pleiotropy can introduce when multiple
instruments are available, but how they perform in the presence of
measurement error remains to be explored. An important advantage that MR
confers over most mediation based analysis is that it can be performed
in two samples, which can considerably improve power and expand the
scope of analysis. However, whether there is a substantive difference in
two sample MR versus one sample MR in how measurement error has an
effect is not yet fully understood. We have also assumed no measurement
error in the genetic instrument, which is not unreasonable given the
strict QC protocols that ensure high quality genotype data is available
to most studies. We have restricted the scope to only exploring
non-differential measurement error and avoided the complications
incurred if measurement error in the exposure and outcome is correlated.
We have also not addressed other issues pertaining to instrumental
variables which are relevant to the question of instrument-exposure
specification. One such problem is exposure misspecification, for
example an instrument could associate with several closely related
putative outcomes, with only one of them actually having a causal effect
on the outcome. This problem has shown to be the case for SNPs
influencing different lipid fractions, for example {[}52,53{]}.

Mediation based network approaches, that go beyond analyses of two
variables, are very well established {[}37{]} and have a number of
extensions that make them valuable tools, including for example network
construction. But because they are predicated on the basic underlying
principles of mediation they are liable to suffer from the same issues
of measurement error. Recent advances in MR methodology, for example
applying MR to genetical genomics {[}54{]}, multivariate MR {[}53{]} and
mediation through MR {[}55--57{]} may offer more robust alternatives for
these more complicated problems.

The overarching result from our simulations is that, regardless of the
method used, inferring the causal direction using an instrument of
unknown biology is highly sensitive to measurement error. With the
presence of measurement error near ubiquitous in most observational
data, and our ability to measure it limited, we argue that it needs to
be central to any consideration of approaches which are used in attempt
to strengthen causal inference, and any putative results should be
accompanied with appropriate sensitivity analysis that assesses their
robustness under varying levels of measurement error.

\newpage

\subsection{Figures}\label{figures}

Figure 1: Gene expression levels (blue blocks) and DNA methylation
levels (green triangles) may be correlated but the causal structure is
unknown. If a SNP (yellow circle) is associated with both DNA
methylation and gene expression levels then it can be used as an
instrument, but there are three basic competing models for these
variables. The causal inference test (CIT) attempts to distinguish
between them. a) Gene expression causes methylation. The left figure
shows that the SNP influences gene expression levels that in turn
influence methylation levels. The right figure shows the directed
acyclic graph that represents this model. Faded symbols represent the
measured values whereas solid symbols represent the true values. b) The
same as in A, except the causal direction is from DNA methylation to
Gene expression. c) A model of confounding, where gene expression and
DNA methylation are not causally related, but the SNP influences them
each through separate pathways or a confounder.

Figure 2: The CIT was performed on simulated variables where the
exposure influenced the outcome and the exposure was instrumented by a
SNP. The test statistic from CIT when testing if the exposure caused the
outcome (the true model) is in red, and the test for the outcome causing
the exposure (false model) is in green. Rows of plots represent the
sample sizes used for the simulations. As measurement imprecision
increases (decreasing values on x-axis) the test statistic for the
incorrect model gets stronger and the test statistic for the correct
model gets weaker.

Figure 3: Outcomes were simulated to be unrelated to the exposure
(bottom plot, showing false positive rates on the y-axis) or causally
influenced by the exposure (top plot, showing true positive rates on the
y-axis) with varying degrees of measurement imprecision applied to the
exposure variable (x axis). Results for MR and CIT were compared for
varying sample sizes (columns of boxes).

Figure 4: a) We can predict the values the MR Steiger test would take
(z-axis) for different potential values of measurement error (x and y
axes), drawn here as the blue surface. When \(\rho_{g,y} > \rho_{g,x}\),
as denoted by the range of values where the blue surface is above the
black plane, those values of measurement error lead to our observed MR
Steiger test inferring the wrong causal direction. Where the blue
surface lies below the black plane, these measurement error values
support the inferred causal direction of X to Y. A measure of
reliability, therefore, is the ratio of the negative and positive
volumes of the total space bound by the blue and black surfaces,
\(R = \frac{V_{z \geq 0}}{ - V_{z < 0} }\). In this case, where
\(\rho_{g,x}^2 = 0.01\) and \(\rho_{x,y}^2 = 0.1\), the \(R = 4.40\),
which means that 4.40 times as much of the possible measurement error
values are in support of the \(x \rightarrow y\) direction of causality
than \(y \rightarrow x\). b) Plots depicting the parameter space in
which the function \(d = cor(x, x_O) - cor(x,y)cor(y, y_O)\) is
negative. When \(d\) is negative the MR Steiger test is liable to infer
the wrong direction of causality. Shaded regions show the parameter
space where \(d\) is negative. The graph shows that for the majority of
the parameter space of the function, \(d\) is positive, especially where
causal relationships are relatively weak.

Figure 5: a) Outcome \(y\) was simulated to be caused by exposure \(x\)
as shown in the graph, with varying degrees of measurement error applied
to both. CIT and MR were used to infer evidence for causality between
the exposure and outcome, and to infer the direction of causality. The
columns of graphs denote intervals for he value of
\(d = \rho_{x, x_o} - \rho_{x,y}\rho_{y,y_o}\), such that when \(d\) is
negative we expect the MR Steiger test to be more likely to be wrong
about the direction of causality. Rows of graphs represent the sample
size used in the simulations. For the CIT method, outcome 1 denoted
evidence for causality with correct model, outcomes 2 or 3 denoted
evidence for causality with incorrect model, and outcome 4 denoted no
evidence for causality. b) As in (a) except the simulated model was
non-causal, and a genetic confounder induces an association between
\(x\) and \(y\). Neither CIT nor MR are able to identify this model, so
any significant associations in MR are deemed to be incorrect, while
outcomes 1 or 2 for the CIT are deemed to be incorrect.

Figure 6: Using 458 putative associations between DNA methylation and
gene expression we used the MR Steiger test to infer the direction of
causality between them. a) The rightmost bar shows the proportion of
associations for each of the two possible causal directions (colour key)
assuming no measurement error in either gene expression or DNA
methylation levels. The proportions change when we assume different
levels of measurement error in gene expression levels (x-axis) or DNA
methylation levels (columns of boxes). If there is systematically higher
measurement error in one platform than the other it will appear to be
less likely to be the causal factor. b) The relationship between the
Pearson correlation between DNA methylation and gene expression levels
(x-axis) and the causal estimate (scaled to be in standard deviation
units, y-axis). c) Distribution of estimated causal effect sizes,
stratified into associations inferred to be due to DNA methylation
causing expression (blue) and expression causing DNA methylation (red).

\subsection{Supplementary information}\label{supplementary-information}

\textbf{S1 Text. The influence of measurement error in the exposure on
mediation-based estimated}

\textbf{S2 Text. Sensitivity analysis for measurement error on the MR
Steiger test}

\textbf{S3 Text. The influence of unmeasured confounding on the
inference of causal directions}

\textbf{S1 Fig. Influence of confounding on CIT.} Illustrative
simulations (\(n = 5000\)) showing the results from CIT analysis under a
model of confounding. Here, the phenotypes \(x\) and \(y\) are not
causally related, but there is a genetic effect and a confounder both
influencing each phenotype. Each point represents a single simulation.
Where power is high (when the absolute values of the \(x\) and \(y\)
axes are large) the CIT returns a significant result (\(p < 0.01\)) when
testing the causal effect of \(x\) on \(y\), and when testing the causal
effect of \(y\) on \(x\).

\textbf{S2 Fig. Influence of confounding on MR Steiger.} Graph
representing the unmeasured confounding parameters that will lead to the
MR Steiger test returning the wrong causal direction. Columns of boxes
represent different signed values of the observational variance
explained between \(x\) and \(y\) (\(R^2_{xy}\)).

\newpage

\subsection{References}\label{references}

\raggedright

\hypertarget{refs}{}
\hypertarget{ref-Phillips1991}{}
1. Phillips AN, Davey Smith G. How independent are ``independent''
effects? relative risk estimation when correlated exposures are measured
imprecisely. Journal of Clinical Epidemiology. Pergamon; 1991;44:
1223--1231.
doi:\href{https://doi.org/10.1016/0895-4356(91)90155-3}{10.1016/0895-4356(91)90155-3}

\hypertarget{ref-DaveySmith2002}{}
2. Davey Smith G, Ebrahim S. Data dredging, bias, or confounding. BMJ.
2002;325: 1437--8.

\hypertarget{ref-DaveySmith2004}{}
3. Davey Smith G, Ebrahim S. Mendelian randomization: prospects,
potentials, and limitations. International journal of epidemiology.
2004;33: 30--42.
doi:\href{https://doi.org/10.1093/ije/dyh132}{10.1093/ije/dyh132}

\hypertarget{ref-Millstein2009}{}
4. Millstein J, Zhang B, Zhu J, Schadt EE. Disentangling molecular
relationships with a causal inference test. BMC genetics. 2009;10: 23.
doi:\href{https://doi.org/10.1186/1471-2156-10-23}{10.1186/1471-2156-10-23}

\hypertarget{ref-Aten2008}{}
5. Aten JE, Fuller TF, Lusis AJ, Horvath S. Using genetic markers to
orient the edges in quantitative trait networks: the NEO software. BMC
systems biology. 2008;2: 34.
doi:\href{https://doi.org/10.1186/1752-0509-2-34}{10.1186/1752-0509-2-34}

\hypertarget{ref-Waszak2015}{}
6. Waszak SM, Delaneau O, Gschwind AR, Kilpinen H, Raghav SK, Witwicki
RM, et al. Variation and genetic control of chromatin architecture in
humans. Cell. Elsevier Inc. 2015;162: 1039--1050.
doi:\href{https://doi.org/10.1016/j.cell.2015.08.001}{10.1016/j.cell.2015.08.001}

\hypertarget{ref-Houle2011}{}
7. Houle D, Pélabon C, Wagner G, Hansen T. Measurement and meaning in
biology. The Quarterly Review of Biology. 2011;86: 3--34. Available:
\url{http://www.jstor.org/stable/10.1086/658408}

\hypertarget{ref-Hernan2009}{}
8. Hernán M a, Cole SR. Invited Commentary: Causal diagrams and
measurement bias. American journal of epidemiology. 2009;170: 959--62;
discussion 963--4.
doi:\href{https://doi.org/10.1093/aje/kwp293}{10.1093/aje/kwp293}

\hypertarget{ref-Harper2013}{}
9. Harper KN, Peters B a, Gamble MV. Batch effects and pathway analysis:
two potential perils in cancer studies involving DNA methylation array
analysis. Cancer epidemiology, biomarkers \& prevention : a publication
of the American Association for Cancer Research, cosponsored by the
American Society of Preventive Oncology. 2013;22: 1052--60.
doi:\href{https://doi.org/10.1158/1055-9965.EPI-13-0114}{10.1158/1055-9965.EPI-13-0114}

\hypertarget{ref-Chen2013a}{}
10. Chen Y-a, Lemire M, Choufani S, Butcher DT, Grafodatskaya D, Zanke
BW, et al. Discovery of cross-reactive probes and polymorphic CpGs in
the Illumina Infinium HumanMethylation450 microarray. Epigenetics :
official journal of the DNA Methylation Society. 2013;8: 203--9.
doi:\href{https://doi.org/10.4161/epi.23470}{10.4161/epi.23470}

\hypertarget{ref-Houseman2012}{}
11. Houseman EA, Accomando WP, Koestler DC, Christensen BC, Marsit CJ,
Nelson HH, et al. DNA methylation arrays as surrogate measures of cell
mixture distribution. BMC bioinformatics. 2012;13: 86.
doi:\href{https://doi.org/10.1186/1471-2105-13-86}{10.1186/1471-2105-13-86}

\hypertarget{ref-Ahima2013}{}
12. Ahima RS, Lazar MA. Physiology. The health risk of obesity--better
metrics imperative. Science (New York, NY). American Association for the
Advancement of Science; 2013;341: 856--8.
doi:\href{https://doi.org/10.1126/science.1241244}{10.1126/science.1241244}

\hypertarget{ref-LeCessie2012}{}
13. Cessie S le, Debeij J, Rosendaal FR, Cannegieter SC, Vandenbroucke
JP. Quantification of bias in direct effects estimates due to different
types of measurement error in the mediator. Epidemiology (Cambridge,
Mass). 2012;23: 551--60.
doi:\href{https://doi.org/10.1097/EDE.0b013e318254f5de}{10.1097/EDE.0b013e318254f5de}

\hypertarget{ref-Blakely2013}{}
14. Blakely T, McKenzie S, Carter K. Misclassification of the mediator
matters when estimating indirect effects. Journal of epidemiology and
community health. 2013;67: 458--66.
doi:\href{https://doi.org/10.1136/jech-2012-201813}{10.1136/jech-2012-201813}

\hypertarget{ref-DaveySmith2003}{}
15. Davey Smith G, Ebrahim S. 'Mendelian randomization': can genetic
epidemiology contribute to understanding environmental determinants of
disease? International Journal of Epidemiology. 2003;32: 1--22.
doi:\href{https://doi.org/10.1093/ije/dyg070}{10.1093/ije/dyg070}

\hypertarget{ref-DaveySmithHemani2014}{}
16. Davey Smith G, Hemani G. Mendelian randomization: genetic anchors
for causal inference in epidemiological studies. Human molecular
genetics. Oxford Univ Press; 2014;23: R89-----R98.
doi:\href{https://doi.org/10.1093/hmg/ddu328}{10.1093/hmg/ddu328}

\hypertarget{ref-Schadt2005}{}
17. Schadt EE, Lamb J, Yang X, Zhu J, Edwards S, GuhaThakurta D, et al.
An integrative genomics approach to infer causal associations between
gene expression and disease. Nature Genetics. 2005;37: 710--717.
doi:\href{https://doi.org/10.1038/ng1589}{10.1038/ng1589}

\hypertarget{ref-Millstein2016}{}
18. Millstein J. cit: Causal Inference Test. R package version 1.9
{[}Internet{]}. 2016. Available:
\url{http://cran.r-project.org/package=cit}

\hypertarget{ref-Koestler2014}{}
19. Koestler DC, Chalise P, Cicek MS, Cunningham JM, Armasu S, Larson
MC, et al. Integrative genomic analysis identifies epigenetic marks that
mediate genetic risk for epithelial ovarian cancer. BMC medical
genomics. BMC Medical Genomics; 2014;7: 8.
doi:\href{https://doi.org/10.1186/1755-8794-7-8}{10.1186/1755-8794-7-8}

\hypertarget{ref-Liu2013}{}
20. Liu Y, Aryee MJ, Padyukov L, Fallin MD, Hesselberg E, Runarsson A,
et al. Epigenome-wide association data implicate DNA methylation as an
intermediary of genetic risk in rheumatoid arthritis. Nature
biotechnology. 2013;31: 142--7.
doi:\href{https://doi.org/10.1038/nbt.2487}{10.1038/nbt.2487}

\hypertarget{ref-Yuan2014}{}
21. Yuan W, Xia Y, Bell CG, Yet I, Ferreira T, Ward KJ, et al. An
integrated epigenomic analysis for type 2 diabetes susceptibility loci
in monozygotic twins. Nature communications. 2014;5: 5719.
doi:\href{https://doi.org/10.1038/ncomms6719}{10.1038/ncomms6719}

\hypertarget{ref-Tang2014}{}
22. Tang Y, Axelsson AS, Spégel P, Andersson LE, Mulder H, Groop LC, et
al. Genotype-based treatment of type 2 diabetes with an
\(\alpha\)2A-adrenergic receptor antagonist. Science translational
medicine. 2014;6: 257ra139.
doi:\href{https://doi.org/10.1126/scitranslmed.3009934}{10.1126/scitranslmed.3009934}

\hypertarget{ref-Hong2015}{}
23. Hong X, Hao K, Ladd-Acosta C, Hansen KD, Tsai H-J, Liu X, et al.
Genome-wide association study identifies peanut allergy-specific loci
and evidence of epigenetic mediation in US children. Nature
communications. 2015;6: 6304.
doi:\href{https://doi.org/10.1038/ncomms7304}{10.1038/ncomms7304}

\hypertarget{ref-Bowden2015}{}
24. Bowden J, Davey Smith G, Burgess S. Mendelian randomization with
invalid instruments: effect estimation and bias detection through Egger
regression. International Journal of Epidemiology. 2015;44: 512--25.

\hypertarget{ref-Bowden2016b}{}
25. Bowden J, Davey Smith G, Haycock PC, Burgess S. Consistent
Estimation in Mendelian Randomization with Some Invalid Instruments
Using a Weighted Median Estimator. Genetic Epidemiology. 2016;40:
304--314.
doi:\href{https://doi.org/10.1002/gepi.21965}{10.1002/gepi.21965}

\hypertarget{ref-Hemani2016}{}
26. Hemani G, Zheng J, Wade KH, Laurin C, Elsworth B, Burgess S, et al.
MR-Base: a platform for systematic causal inference across the phenome
using billions of genetic associations. BioRxiv. 2016;10.1101/07.

\hypertarget{ref-Timpson2011}{}
27. Timpson NJ, Nordestgaard BG, Harbord RM, Zacho J, Frayling TM,
Tybjærg-Hansen a, et al. C-reactive protein levels and body mass index:
elucidating direction of causation through reciprocal Mendelian
randomization. International journal of obesity (2005). 2011;35: 300--8.
doi:\href{https://doi.org/10.1038/ijo.2010.137}{10.1038/ijo.2010.137}

\hypertarget{ref-Richmond2014}{}
28. Richmond RC, Davey Smith G, Ness AR, Hoed M den, McMahon G, Timpson
NJ. Assessing Causality in the Association between Child Adiposity and
Physical Activity Levels: A Mendelian Randomization Analysis. Ludwig DS,
editor. PLoS Medicine. 2014;11: e1001618.
doi:\href{https://doi.org/10.1371/journal.pmed.1001618}{10.1371/journal.pmed.1001618}

\hypertarget{ref-Claussnitzer2015}{}
29. Claussnitzer M, Dankel SN, Kim K-H, Quon G, Meuleman W, Haugen C, et
al. FTO Obesity Variant Circuitry and Adipocyte Browning in Humans. The
New England journal of medicine. 2015;373: 895--907.
doi:\href{https://doi.org/10.1056/NEJMoa1502214}{10.1056/NEJMoa1502214}

\hypertarget{ref-Pierce2013}{}
30. Pierce BL, Burgess S. Efficient design for Mendelian randomization
studies: subsample and 2-sample instrumental variable estimators.
American journal of epidemiology. 2013;178: 1177--84.
doi:\href{https://doi.org/10.1093/aje/kwt084}{10.1093/aje/kwt084}

\hypertarget{ref-Hernan2004}{}
31. Hernán MA, Hernández-Díaz S, Robins JM. A structural approach to
selection bias. Epidemiology (Cambridge, Mass). 2004;15: 615--25.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/15308962}

\hypertarget{ref-Pierce2012}{}
32. Pierce BL, VanderWeele TJ. The effect of non-differential
measurement error on bias, precision and power in Mendelian
randomization studies. International Journal of Epidemiology. Oxford
University Press; 2012;41: 1383--1393.
doi:\href{https://doi.org/10.1093/ije/dys141}{10.1093/ije/dys141}

\hypertarget{ref-Ashenfelter1994}{}
33. Ashenfelter O, Krueger AB. Estimates of the Economic Return to
Schooling from a New Sample of Twins. The American Economic Review.
1994;84: 1157--1173.
doi:\href{https://doi.org/10.1016/S0272-7757(98)00038-7}{10.1016/S0272-7757(98)00038-7}

\hypertarget{ref-Nagarajan2013}{}
34. Nagarajan R, Scutari M. Impact of noise on molecular network
inference. PloS one. 2013;8: e80735.
doi:\href{https://doi.org/10.1371/journal.pone.0080735}{10.1371/journal.pone.0080735}

\hypertarget{ref-Shpitser2010}{}
35. Shpitser I, VanderWeele T, Robins J. On the validity of covariate
adjustment for estimating causal effects. Proceedings of the Twenty
Sixth Conference on Uncertainty in Artificial Intelligence (UAI-10).
2010; 527--536.

\hypertarget{ref-Wang2015}{}
36. Wang L, Michoel T. Detection of regulator genes and eQTLs in gene
networks. arXiv. 2015;arXiv:1512. Available:
\url{http://arxiv.org/abs/1512.05574}

\hypertarget{ref-Lagani2015}{}
37. Lagani V, Triantafillou S, Ball G, Tegner J, Tsamardinos I.
Probabilistic Computational Causal Discovery for Systems Biology.
Uncertainty in biology: A computational modeling approach. Springer;
2015. p. 47. Available:
\href{https://books.google.com/books?id=8SLUCgAAQBAJ\%7B/\&\%7Dpgis=1}{https://books.google.com/books?id=8SLUCgAAQBAJ\{\textbackslash{}\&\}pgis=1}

\hypertarget{ref-Lawlor2017}{}
38. Lawlor DA, Tilling K, Davey Smith G. Triangulation in aetiological
epidemiology. International Journal of Epidemiology. 2017;19: dyw314.
doi:\href{https://doi.org/10.1093/ije/dyw314}{10.1093/ije/dyw314}

\hypertarget{ref-Sterne2001}{}
39. Sterne JAC, Smith GD. Sifting the evidence---what's wrong with
significance tests? BMJ. 2001;322: 226--231.

\hypertarget{ref-Henningsen2007}{}
40. Henningsen A, Hamann JD. systemfit : A Package for Estimating
Systems of Simultaneous Equations in R. Journal of Statistical Software.
2007;23: 1--40.
doi:\href{https://doi.org/10.18637/jss.v023.i04}{10.18637/jss.v023.i04}

\hypertarget{ref-Steiger1980}{}
41. Steiger JH. Tests for comparing elements of a correlation matrix.
Psychological Bulletin. 1980;87: 245--251.
doi:\href{https://doi.org/10.1037/0033-2909.87.2.245}{10.1037/0033-2909.87.2.245}

\hypertarget{ref-Revelle2015}{}
42. Revelle W. psych: Procedures for Psychological, Psychometric, and
Personality Research {[}Internet{]}. Evanston, Illinois: Northwestern
University; 2015. Available:
\url{http://cran.r-project.org/package=psych}

\hypertarget{ref-Shakhbazov2016}{}
43. Shakhbazov K, Powell JE, Hemani G, Henders AK, Martin NG, Visscher
PM, et al. Shared genetic control of expression and methylation in
peripheral blood. BMC genomics. BioMed Central; 2016;17: 278.
doi:\href{https://doi.org/10.1186/s12864-016-2498-4}{10.1186/s12864-016-2498-4}

\hypertarget{ref-RCoreTeam2015}{}
44. R Core Team. R: A Language and Environment for Statistical Computing
{[}Internet{]}. Vienna, Austria: R Foundation for Statistical Computing;
2015. Available: \url{https://www.r-project.org/}

\hypertarget{ref-Bird2002}{}
45. Bird A. DNA methylation patterns and epigenetic memory. Genes \&
development. 2002;16: 6--21.
doi:\href{https://doi.org/10.1101/gad.947102}{10.1101/gad.947102}

\hypertarget{ref-Cole2014}{}
46. Cole DA, Preacher KJ. Manifest Variable Path Analysis: Potentially
Serious and Misleading Consequences Due to Uncorrected Measurement
Error. Psychological Methods. 2014;19: 300--315.
doi:\href{https://doi.org/10.1037/a0033805}{10.1037/a0033805}

\hypertarget{ref-Bose2014}{}
47. Bose M, Wu C, Pankow JS, Demerath EW, Bressler J, Fornage M, et al.
Evaluation of microarray-based DNA methylation measurement using
technical replicates: the Atherosclerosis Risk In Communities (ARIC)
Study. BMC Bioinformatics. 2014;15: 312.
doi:\href{https://doi.org/10.1186/1471-2105-15-312}{10.1186/1471-2105-15-312}

\hypertarget{ref-Bryant2011}{}
48. Bryant PA, Smyth GK, Robins-Browne R, Curtis N, Novak J, Sladek R,
et al. Technical Variability Is Greater than Biological Variability in a
Microarray Experiment but Both Are Outweighed by Changes Induced by
Stimulation. Khodursky AB, editor. PLoS ONE. Public Library of Science;
2011;6: e19556.
doi:\href{https://doi.org/10.1371/journal.pone.0019556}{10.1371/journal.pone.0019556}

\hypertarget{ref-Zhu2016}{}
49. Zhu Z, Zhang F, Hu H, Bakshi A, Robinson MR, Powell JE, et al.
Integration of summary data from GWAS and eQTL studies predicts complex
trait gene targets. Nature Genetics. Nature Research; 2016;48: 481--487.
doi:\href{https://doi.org/10.1038/ng.3538}{10.1038/ng.3538}

\hypertarget{ref-Mancuso2017}{}
50. Mancuso N, Shi H, Goddard P, Kichaev G, Gusev A, Pasaniuc B.
Integrating Gene Expression with Summary Association Statistics to
Identify Genes Associated with 30 Complex Traits. The American Journal
of Human Genetics. 2017;100: 473--487.
doi:\href{https://doi.org/10.1016/j.ajhg.2017.01.031}{10.1016/j.ajhg.2017.01.031}

\hypertarget{ref-Lee2013c}{}
51. Lee SH, Wray NR. Novel genetic analysis for case-control genome-wide
association studies: quantification of power and genomic prediction
accuracy. PLoS One. 2013;8: e71494.
doi:\href{https://doi.org/10.1371/journal.pone.0071494\%20\%5Bdoi\%5D/rPONE-D-13-18776\%20\%5Bpii\%5D}{10.1371/journal.pone.0071494 {[}doi{]}\textbackslash{}rPONE-D-13-18776 {[}pii{]}}

\hypertarget{ref-Do2013}{}
52. Do R, Willer CJ, Schmidt EM, Sengupta S, Gao C, Peloso GM, et al.
Common variants associated with plasma triglycerides and risk for
coronary artery disease. Nature Genetics. Nature Research; 2013;45:
1345--1352. doi:\href{https://doi.org/10.1038/ng.2795}{10.1038/ng.2795}

\hypertarget{ref-Burgess2014a}{}
53. Burgess S, Freitag DF, Khan H, Gorman DN, Thompson SG. Using
multivariable Mendelian randomization to disentangle the causal effects
of lipid fractions. PloS one. Public Library of Science; 2014;9:
e108891.
doi:\href{https://doi.org/10.1371/journal.pone.0108891}{10.1371/journal.pone.0108891}

\hypertarget{ref-Relton2012}{}
54. Relton CL, Davey Smith G. Two-step epigenetic Mendelian
randomization: a strategy for establishing the causal role of epigenetic
processes in pathways to disease. International journal of epidemiology.
2012;41: 161--76.
doi:\href{https://doi.org/10.1093/ije/dyr233}{10.1093/ije/dyr233}

\hypertarget{ref-Varbo2015}{}
55. Varbo A, Benn M, Smith GD, Timpson NJ, Tybjaerg-Hansen A,
Nordestgaard BG. Remnant cholesterol, low-density lipoprotein
cholesterol, and blood pressure as mediators from obesity to ischemic
heart disease. Circulation research. 2015;116: 665--73.
doi:\href{https://doi.org/10.1161/CIRCRESAHA.116.304846}{10.1161/CIRCRESAHA.116.304846}

\hypertarget{ref-Burgess2015}{}
56. Burgess S, Daniel RM, Butterworth AS, Thompson SG. Network Mendelian
randomization: using genetic variants as instrumental variables to
investigate mediation in causal pathways. International journal of
epidemiology. 2015;44: 484--95.
doi:\href{https://doi.org/10.1093/ije/dyu176}{10.1093/ije/dyu176}

\hypertarget{ref-Richmond2016}{}
57. Richmond RC, Hemani G, Tilling K, Davey Smith G, Relton CL.
Challenges and novel approaches for investigating molecular mediation.
Human molecular genetics. Oxford University Press; 2016;25: R149--R156.
doi:\href{https://doi.org/10.1093/hmg/ddw197}{10.1093/hmg/ddw197}

\end{document}