### 2017-10-24 docs/s3text.Rmd

---
title: "Supplementary materials: Orienting the causal relationship between imprecisely measured traits using GWAS summary data"
output:
  pdf_document:
    fig_caption: no
    keep_tex: true
  word_document: default
---

## S3 Text. The influence of unmeasured confounding on the inference of causal directions

We have assumed no unmeasured confounding in these simulations. Unmeasured confounding will however have potentially large influences on mediation-based methods for inferring causal directions, and can also adversely influence the estimate of the causal direction for the Steiger test.

### Unmeasured confounding in mediation

Including an unmeasured confounder, $u$, after ignoring intercept terms the exposure $x$ and outcome $y$ variables can be modelled as

$$
\begin{aligned}
y & = \beta_x x + \beta_{uy} u + \epsilon_x \\
x & = \beta_g g + \beta_{ux} u + \epsilon_g
\end{aligned}
$$

The observational estimate of the causal effect of $x$ on $y$, $\hat{\beta}_x$ is obtained from

$$
\begin{aligned}
\hat{\beta}_x & = cov(x, y) / var(x) \\
& = \frac{\beta_g^2 \beta_x var(g) + \beta_{ux}^2 \beta_x var(u) + \beta_x var(\epsilon_g)} {\beta_g^2 var(g) + \beta_{ux}^2 var(u) + var(\epsilon_g)}
\end{aligned}
$$

From this it is clear that $\beta_x$ and $\hat{\beta}_x$ will differ when both $\beta_{uy}$ and $\beta_{ux}$ are non-zero. Relating to mediation, where we attempt to test if $g$ associates with $y$ after adjusting $y$ for $x$, such that

$$
\hat{y}^* = \hat{\beta}_x x
$$

and

$$
\begin{aligned}
cov(g, y - \hat{y}^*) & = cov(g, \beta_x x + \beta_{uy} u + \epsilon_x - \hat{\beta}_x x) \\
& = cov(g, (\beta_x - \hat{\beta}_x)(\beta_g g + \beta_ux u + \epsilon_x)) \\
& = (\beta_x - \hat{\beta}_x) var(g)
\end{aligned}
$$

should any amount of unmeasured confounding exist, therefore, there will remain an association between $g$ and $y|x$, which will introduce errors in inferring causal directions.

### Unmeasured confounding in the MR Steiger test

Similarly, we can investigate the extent to which unmeasured confounding will lead to the wrong causal direction between $x$ and $y$ using the MR Steiger test, evaluating the liability for the inequality $cor(g,x)^2 > cor(g,y)^2$ being incorrect. After some algebra

$$
cor(g,x)^2 = \frac{\beta_g^2}{\beta_g^2var(g) + \beta_{ux}^2 var(u) + var(\epsilon_x)}
$$

and

$$
cor(g,y)^2 = \frac{\beta_x^2 \beta_g^2 var(g)^2} {\hat{\beta}_x^2 \beta_g^2 var(g) + \hat{\beta}_x^2 \beta_{ux}^2 var(u) + \beta_{uy}^2 var(u) + var(\epsilon_y)}
$$

S2 fig shows the relationship between the magnitude of the correlations between $x$, $y$ and the confounder $u$ for a range of $\beta_{xy} = (-2,2)$, $\beta_{gx} = 0.1$ and a range of confounder effects. The pattern of results were similar for different values of $\beta_{gx}$. We note that in most cases for the parameter values explored, where the observational absolute $\hat{R}^2_{xy}$ is less than 0.2, unmeasured confounding will not incur the wrong causal direction in the MR Steiger test.